### 2025-10-27 Text/Thesis/Chapters/Chapter4.qmd
final version thesis

Figure&nbsp;\ref{fig:sim_Phantom_1} presents the structural design of the synthetic phantom with dimensions of
$500\times500$ pixels, proposed by  Gomez et al.&nbsp;[-@Gomez2017] as a tool to assess the performance of speckle-reduction filters. This layout serves as a base template to generate the simulated images shown in Figures&nbsp;\ref{fig:sim_Phantom_2}–\ref{fig:sim_Phantom_3}.

Each region is filled with data simulated from the $\mathcal{G}^0_I$ distribution&nbsp;\eqref{E:gi02}, using different combinations of the roughness parameter $\alpha$ and the mean $\mu$, as indicated by the labels in each quadrant of the image. Light regions correspond to textured observations (heterogeneous), while darker regions represent textureless areas (homogeneous).
```{=latex}
\begin{figure}[H]
  \centering
  \hspace{0.5em}
  \begin{subfigure}[b]{0.37\textwidth}
    \centering
    \includegraphics[width=\linewidth]{Figures/Phantom1}
    \caption{Phantom.}
    \label{fig:sim_Phantom_1}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.50\textwidth}
    \centering
    \includegraphics[width=\linewidth]{Figures/Phantom_labels}
    \caption{Simulated image, varying $\alpha$ and $\mu$, with $L=5$.}
    \label{fig:sim_Phantom_2}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.50\textwidth}
    \centering
    \includegraphics[width=\linewidth]{Figures/Phantom4_L9}
    \caption{Simulated image, varying $\alpha$ and $\mu$, with $L=9$.}
    \label{fig:sim_Phantom_3}
  \end{subfigure}
  \caption{Synthetic dataset.}
  \label{fig:sim_Phantom}
\end{figure}
```
The first simulated image uses $L = 5$, and the second uses $L = 9$, to evaluate how the test statistics behave as the number of looks increases and speckle noise is reduced.

The $\alpha$ parameter of the $G_I^0$ distribution is essential for
interpreting texture characteristics. Values near zero greater than
$-3$ suggest extremely textured targets, such as urban
zones&nbsp;[@Frery2019a]. As the value decreases, it indicates regions
with moderate texture (in the $\left[-6,-3\right]$ region), related to
forest zones, while values below $-6$ correspond to textureless
regions, such as pasture, agricultural fields, and water
bodies&nbsp;[@Neto2023].

We applied the three test statistics, $S_{\widetilde{H}_{\text{AO}}}$,  $S_{\widetilde{R}_{\lambda}}$, and
$S_{\widetilde{T}_{\lambda}}$, to the simulated images using local sliding windows of size $7\times 7$. Figures&nbsp;\ref{fig:sim1r} and&nbsp;\ref{fig:sim2t} show the $p$-value maps (top row) and the corresponding binary maps (bottom row) for two simulations with $L=5$ and $L=9$, respectively. The $p$-value maps (Figures&nbsp;\ref{fig:sim1a}–\ref{fig:sim1c} and \ref{fig:sim2a}–\ref{fig:sim2c}) are displayed as gradient color maps, where dark blue tones (close to $0$) indicate strong evidence of heterogeneity, and lighter tones (toward red, near $1$) correspond to more homogeneous areas.

The binary maps (Figures&nbsp;\ref{fig:sim1d}–\ref{fig:sim1f} and \ref{fig:sim2d}–\ref{fig:sim2f}) simplify these results using a threshold of $0.05$: pixels corresponding to $p$-values below this threshold are shown in black, suggesting statistically significant deviations and thus indicating heterogeneity. In contrast, pixels with $p \geq 0.05$ are displayed in white, representing no evidence to reject the null hypothesis of homogeneity.

```{=latex}
\begin{figure}[hbt]
    \centering

    % Primera fila
    \begin{subfigure}{0.3\textwidth}
        \includegraphics[width=\linewidth]{Figures/Phantom_4_z_b200_w7-AO.png}
        \caption{$p$-value map $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:sim1a}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.3\textwidth}
        \includegraphics[width=\linewidth]{Figures/Phantom_4_z_0.9_b200_w7-renyi.png}
        \caption{$p$-value map $\small{S_{\widetilde{R}_{\lambda}}}$}
        \label{fig:sim1b}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.3\textwidth}
        \includegraphics[width=\linewidth]{Figures/Phantom_4_z_0.85_b200_w7-tsallis77.png}
        \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$}
        \label{fig:sim1c}
    \end{subfigure}

    \vspace{1em} % Espacio entre filas

    % Segunda fila
    \hspace{-2.5em}
    \begin{subfigure}{0.25\textwidth}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom_4_z_b200_w7-AO.png}
        \caption{Binary map $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:sim1d}
    \end{subfigure}
    \hspace{2.8em}
    \begin{subfigure}{0.25\textwidth}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom_4_z_0.9_b200_w7-renyi.png}
        \caption{Binary map  $\small{S_{\widetilde{R}_{\lambda}}}$}
        \label{fig:sim1e}
    \end{subfigure}
     \hspace{3.1em}
    \begin{subfigure}{0.25\textwidth}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom_4_z_0.85_b200_w7-tsallis.png}
        \caption{Binary map  $\small{S_{\widetilde{T}_{\lambda}}}$}
        \label{fig:sim1f}
    \end{subfigure}
    \caption{Results on simulated data ($L = 5$) using $S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$, and $S_{\widetilde{T}_{\lambda}}$. Top: $p$-value heatmaps. Bottom: binary maps ($5\%$ level).}
%\caption{Comparison of test statistics $S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$, and $S_{\widetilde{T}_{\lambda}}$ on a simulated image with $L = 5$. Top row: heatmaps of $p$-values computed using a $7 \times 7$ sliding window. Bottom row: binary maps at the 5\% significance level, where black pixels indicate rejection of the null hypothesis of homogeneity.}
    \label{fig:sim1r}
\end{figure}
```

```{=latex}
\begin{figure}[hbt]
    \centering

    % Primera fila
    \begin{subfigure}{0.3\textwidth}
        \includegraphics[width=\linewidth]{Figures/Phantom4_L9_1_8_b100_w7-AO-H.png}
        \caption{$p$-value map $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:sim2a}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.3\textwidth}
        \includegraphics[width=\linewidth]{Figures/Phantom4_L9_1_8_b100_w7-09-renyi.png}
        \caption{$p$-value map $\small{S_{\widetilde{R}_{\lambda}}}$}
        \label{fig:sim2b}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.3\textwidth}
        \includegraphics[width=\linewidth]{Figures/Phantom4_L9_1_8_0.85_b100_w7-tsallis.png}
        \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$}
        \label{fig:sim2c}
    \end{subfigure}

    \vspace{1em} % Espacio entre filas

    % Segunda fila
    \hspace{-1.6em}
    \begin{subfigure}{0.25\textwidth}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom4_L9_1_8_b100_w7-AO.png}
        \caption{Binary map $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:sim2d}
    \end{subfigure}
    \hspace{3.0em}
    \begin{subfigure}{0.25\textwidth}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom4_L9_1_8_b100_w7-09-renyi.png}
        \caption{Binary map  $\small{S_{\widetilde{R}_{\lambda}}}$}
        \label{fig:sim2e}
    \end{subfigure}
     \hspace{3.1em}
    \begin{subfigure}{0.25\textwidth}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom4_L9_1_8_0.85_b100_w7-tsallis.png}
        \caption{Binary map  $\small{S_{\widetilde{T}_{\lambda}}}$}
        \label{fig:sim2f}
    \end{subfigure}
    \caption{Results on simulated data ($L = 9$) using $S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$, and $S_{\widetilde{T}_{\lambda}}$. Top: $p$-value heatmaps. Bottom: binary maps ($5\%$ level).}
   %  \caption{Comparison of test statistics $S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$, and $S_{\widetilde{T}_{\lambda}}$ on a simulated image with $L = 9$. Top row: heatmaps of $p$-values computed using a $7 \times 7$ sliding window. Bottom row: binary maps at the 5\% significance level, where black pixels indicate rejection of the null hypothesis of homogeneity.}
    \label{fig:sim2t}
\end{figure}
```

When the number of looks increases to $L=9$ (Figure&nbsp;\ref{fig:sim2t}), the maps become cleaner due to reduced speckle.
All tests are able to detect zones with different levels of texture.
However, some differences can be observed. The statistic $S_{\widetilde{R}_{\lambda}}$, based on Rényi entropy, is very sensitive to changes in texture and often highlights small variations and detailed patterns of heterogeneity. In contrast, $S_{\widetilde{T}_{\lambda}}$, based on Tsallis entropy, tends to avoid false detections in uniform areas and produces cleaner results. This is due to the underlying formulation: Rényi entropy emphasizes variability, making it effective in identifying heterogeneity; while Tsallis entropy is more robust to noise, which helps preserve homogeneous regions.

@@

### SAR dataset and preprocessing {#sec:SAR-data}

We evaluated the proposed test statistics using a set of eight SAR images acquired from different radar missions. These images cover diverse locations: the surroundings of London (United Kingdom), San Francisco (USA), Rotterdam (Netherlands), the outskirts of Munich and the city of New Orleans (USA), Dublin Port (Ireland), the coast of Jalisco (Mexico), and the area around Foggia (Italy) as shown in Figures&nbsp;\ref{fig:london-sar}-\ref{fig:fo-sar}. The selected scenes include a variety of land types, such as urban zones, mountainous areas, agricultural regions, and water bodies. This diversity allows us to assess the performance of the proposed statistical tests under different texture conditions.

To help interpret the SAR images, we also included corresponding optical images (Figures&nbsp;\ref{fig:london-optical}-\ref{fig:fo-optical}) obtained from Sentinel-2 Level-2A products. These optical scenes were acquired on dates close to those of the SAR acquisitions to ensure visual consistency. The optical data were  downloaded from the Copernicus Data Space platform\footnote{\url{https://dataspace.copernicus.eu/}}.

The SAR images come from different missions: TanDEM-X, PAZ, and ICEYE (all operating in the X-band), UAVSAR (L-band), and Sentinel-1B (C-band). The two UAVSAR images were obtained from the Alaska Satellite Facility (ASF) platform\footnote{\url{https://search.asf.alaska.edu/}}, which provides access to airborne and satellite SAR data. The remaining images were downloaded from the European Space Agency’s Third Party Missions (ESA TPM) data service\footnote{\url{https://tpm-ds.eo.esa.int/collections/}}, which offers SAR products from various external missions operated by international partners and commercial providers. All images are single-polarization (HH or VV), and their acquisition parameters vary in resolution, scene size, and number of looks.

Table&nbsp;\ref{tab:table_param} summarizes the main characteristics of each SAR image, including the spatial resolution, acquisition date, and number of looks.

@@

SAR_data <- data.frame(
  Site = c("London", "San Francisco",  "Rotterdam", "Munich","New Orleans", "Dublin","Coast of Jalisco", "Foggia" ),
  Mission = c("TanDEM-X", "PAZ", "TanDEM-X","UAVSAR", "UAVSAR", "TanDEM-X", "Sentinel-1B", "Iceye"),
  Band = c("X",  "X","X","L","L", "X", "C", "X"),
  Polarization = c("HH", "HH", "HH","HH","HH","HH", "VV", "VV"),
  Size = c("$2000\\times2000$", "$3300\\times3300$",  "$1200\\times1200$", "$1024\\times1024$","$1400\\times1400$", "$1100\\times1100$","$512\\times512$","$1900\\times1900$"),
  L = c(1, 1, 9, 12, 12,  16,18, 24),
  Resol = c("$0.99/0.99$", "$1.84/1.84$",   "$2.5/2.5$", "$4.9/7.2$","$4.9/7.2$", "$1.35/1.35$","$40/40$", "$2.08/2.08$"),
  Date = c("12-11-2021","30-07-2020","06-12-2018", "16-04-2015","13-03-2016", "03-09-2017","29-08-2021","28-03-2024")
)

@@

We applied the three bootstrap statistics
$S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$ and
$S_{\widetilde{T}_{\lambda}}$ to eight intensity images as shown Figures&nbsp;\ref{fig:london-sar}-\ref{fig:fo-sar} using $7\times 7$ local sliding windows. Each figure is organised in three rows:
the first row displays the SAR scene and a matching Sentinel-2
optical image, the second row shows heat-maps of the
$p$-values produced by the three tests, and the third row presents the
binary decisions at the $5\%$ significance level
(black $\,=\,$ heterogeneous, white $\,=\,$ homogeneous).

Across all scenes the three tests identify the main textured elements
(urban blocks, harbor facilities, roads, crop boundaries, mountainous areas, and river
banks).  In the $p$-value heat-maps dark-blue tones
($p\!<\!0.05$) cluster over these features, whereas red tones
($p\!\approx\!1$) dominate smooth areas such as open water, uniform
pasture or crop areas.  The corresponding binary maps make this pattern explicit, delineating heterogeneous areas with a black mask.

The London and San Francisco images, acquired with $L=1$, exhibit strong speckle, making heterogeneity detection more difficult. 
Among the three tests, the Rényi-based statistic $S_{\widetilde{R}{\lambda}}$ offers the most informative response: its $p$-value maps show dark-blue tones over heterogeneous regions and strong red tones ($p \approx 1$) over clearly homogeneous areas such as water bodies and smooth crop zones.
In contrast, the Shannon and Tsallis-based tests produce $p$-value maps with softer, intermediate colors, which makes it harder to clearly distinguish between textured and smooth areas. Although the binary maps look generally similar, $S_{\widetilde{R}_{\lambda}}$ gives more precise and reliable detections, even under strong speckle.

As the number of looks increases (Rotterdam, Munich and
the remaining images) speckle is reduced and all tests
improve, but differences persist:
$S_{\widetilde{R}_{\lambda}}$ keeps the highest sensitivity to
texture changes, whereas $S_{\widetilde{T}_{\lambda}}$ produces the
cleanest white regions, indicating a stronger ability to preserve
truly homogeneous surfaces.  $\:S_{\widetilde{H}_{\text{AO}}}$ shows an
intermediate behaviour.

These results mirror the properties already observed with simulated
data.  The Rényi-based statistic reacts quickly to any local
fluctuation, making it very effective for spotting heterogeneity even
under severe speckle.  The Tsallis-based test is
more conservative: it suppresses small, noisy variations and therefore
preserves large homogeneous patches, while still detecting clear
textural boundaries.  The Shannon-based test offers a balanced
compromise but is less robust when $L$ is very low.

@@

```{=latex}
\begin{figure}[H]
    \centering

    % Primera fila: 2 imágenes
    \begin{subfigure}{0.3\textwidth}
        \includegraphics[width=\linewidth]{Figures/london_2000.png}
        \caption{SAR image}
        \label{fig:london-sar}
    \end{subfigure}
    \hspace{2em}
    \begin{subfigure}{0.29\textwidth}
        \includegraphics[width=\linewidth]{Figures/london_optical_.png}
        \caption{Optical image}
        \label{fig:london-optical}
    \end{subfigure}

    \vspace{1.2em}

    \begin{subfigure}{0.31\textwidth}
        \includegraphics[width=\linewidth]{Figures/p-values_shannon-london_H1.png}
        \caption{$p$-value map for $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:london-shann1}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.31\textwidth}
        \includegraphics[width=\linewidth]{Figures/p-values_renyi-london_H1.png}
        \caption{$p$-value map for $\small{S_{\widetilde{R}_{\lambda}}}$}
        \label{fig:london-shann-renyi1}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.31\textwidth}
        \includegraphics[width=\linewidth]{Figures/p-values_tsallis-london_H1_new.png} % <--
        \caption{$p$-value map for $\small{S_{\widetilde{T}_{\lambda}}}$}
        \label{fig:london-tsallis1}
    \end{subfigure}

    \vspace{1.2em}

>>>>

Figure&nbsp;\ref{fig:sim_Phantom_1} presents the structural design of the synthetic phantom with dimensions of
$500\times500$ pixels, proposed by  Gomez et al.&nbsp;[-@Gomez2017] as a tool to assess the performance of speckle-reduction filters. This layout serves as a base template to generate the simulated images shown in Figures&nbsp;\ref{fig:sim_Phantom_2}–\ref{fig:sim_Phantom_3}.
Each region is filled with data simulated from the $\mathcal{G}^0_I$ distribution&nbsp;\eqref{E:gi02}, using different combinations of the roughness parameter $\alpha$ and the mean $\mu$, as indicated by the labels in each quadrant of the image. Light regions correspond to textured observations (heterogeneous), while darker regions represent textureless areas (homogeneous).
```{=latex}
\begin{figure}[H]
  \centering
  \caption{Synthetic dataset.}
  \label{fig:sim_Phantom}
  \hspace{2em}
  \begin{subfigure}[b]{0.34\textwidth}
    \centering
    \caption{Phantom.}
    \label{fig:sim_Phantom_1}
    \includegraphics[width=\linewidth]{Figures/Phantom1}
    \end{subfigure}
  \vspace{1.2em} 
  %\hspace{0.48em}
  \begin{subfigure}[b]{0.43\textwidth}
    \centering
    \caption{Simulated image, varying $\alpha$ and $\mu$, with $L=5$.}
    \label{fig:sim_Phantom_2}
    \includegraphics[width=\linewidth]{Figures/Phantom_labels}
      \end{subfigure}
  \hfill
  %\hspace{0.48em}
  \begin{subfigure}[b]{0.43\textwidth}
    \centering
    \caption{Simulated image, varying $\alpha$ and $\mu$, with $L=9$.}
    \label{fig:sim_Phantom_3}
    \includegraphics[width=\linewidth]{Figures/Phantom4_L9}
      \end{subfigure}
  %\caption*{\footnotesize\textbf{Source:} The author (2025)}
\end{figure}
```
The first simulated image uses $L = 5$, and the second uses $L = 9$, to evaluate how the test statistics behave as the number of looks increases and speckle noise is reduced.

The $\alpha$ parameter of the $\mathcal{G}_I^0$ distribution is essential for
interpreting texture characteristics. Values near zero greater than
$-3$ suggest extremely textured targets, such as urban
zones&nbsp;[@Frery2019a]. As the value decreases, it indicates regions
with moderate texture (in the $\left[-6,-3\right]$ region), related to
forest zones, while values below $-6$ correspond to textureless
regions, such as pasture, agricultural fields, and water
bodies&nbsp;[@Neto2023].

We evaluated the three test statistics $S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$, and $S_{\widetilde{T}_{\lambda}}$ on simulated images using fixed sliding windows of size $7\times7$ and, for Tsallis only, the adaptive windowing strategy with side lengths from $5\times5$ ($N_{\min}=5$) to $11\times11$ ($N_{\max}=11$) and smoothing parameter $\eta=3$. Figures&nbsp;\ref{fig:sim1t} and \ref{fig:sim2t} present, respectively for $L=5$ and $L=9$, the $p$-value maps (top row) and the corresponding binary decision maps (bottom row).

The $p$-value maps (Figures&nbsp;\ref{fig:sim1a}–\ref{fig:sim1d} and \ref{fig:sim2a}–\ref{fig:sim2d}) are shown with a continuous color scale: dark red ($p\!\approx\!0$) indicates strong evidence of heterogeneity, while dark blue ($p\!\approx\!1$) corresponds to homogeneous areas. Intermediate colors follow a smooth progression orange ($p\!\approx\!0.2$–$0.4$) for transitional textures and yellow–green to cyan ($p\!\approx\!0.4$–$0.8$) for increasingly near-homogeneous regions.

The binary maps (Figures&nbsp;\ref{fig:sim1e}–\ref{fig:sim1h} and \ref{fig:sim2e}–\ref{fig:sim2h}) simplify these results using a threshold of $0.05$. Pixels with $p$-values below this threshold are shown in black, indicating statistically significant deviations and, therefore, heterogeneity. In contrast, pixels with $p \geq 0.05$ are displayed in white, showing no evidence to reject the null hypothesis of homogeneity.

When the number of looks increases to $L=9$ (Figure&nbsp;\ref{fig:sim2t}), the maps become cleaner due to reduced speckle. All tests successfully detect areas with different texture levels, although some differences remain. The Rényi-based statistic $S_{\widetilde{R}_{\lambda}}$ is more sensitive to texture variations, highlighting fine patterns of heterogeneity, while the Tsallis-based $S_{\widetilde{T}_{\lambda}}$ produces cleaner results in uniform areas by reducing false detections. This behavior reflects their formulations: Rényi entropy emphasizes variability, whereas Tsallis entropy is more robust to noise, helping to preserve homogeneous regions.

The adaptive Tsallis version further improves visual quality, larger windows in homogeneous areas yield smoother estimates, and smaller windows in heterogeneous regions enhance detail preservation.

```{=latex}
\begingroup
\setlength{\textfloatsep}{6pt} % menos espacio entre texto y figuras
\setlength{\floatsep}{6pt} % menos espacio entre figuras
\setlength{\intextsep}{6pt}

\begin{figure}[hbtp]
    \centering
 \caption{Results on simulated data ($L = 5$) using $S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$, and $S_{\widetilde{T}_{\lambda}}$. (a-c) $p$-value maps with a fixed $7\times7$ window.
(d) $p$-value map with adaptive windows ($W_{\min}=5$, $W_{\max}=11$) for
$\eta=3.0$. (e-h) Binary decision maps at 0.05 significance level.}
    \label{fig:sim1t}
    % Primera fila
    \begin{subfigure}{0.24\textwidth}
    \caption{$p$-value map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:sim1a}
        \includegraphics[width=\linewidth]{Figures/p-values_L5_4_z_b100_w7-AO.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:sim1b}
        \includegraphics[width=\linewidth]{Figures/p-values_L5_4_z_b100_w7-renyi.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
    \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:sim1c}
        \includegraphics[width=\linewidth]{Figures/p-values_L5_4_z_b100_w7-tsallis.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
    \caption{$p$-value map $\tiny{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
        \label{fig:sim1d}
        \includegraphics[width=\linewidth]{Figures/p-values_L5_Tsallis_adaptive.png}
            \end{subfigure}
    \vspace{1em} % Espacio entre filas

    % Segunda fila
    \hspace{-0.5em}
    \begin{subfigure}{0.23\textwidth}
    \caption{Binary map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:sim1e}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom_4_z_b200_w7-AO.png}
            \end{subfigure}
    \hspace{0.5em}
    \begin{subfigure}{0.23\textwidth}
    \caption{Binary map  $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:sim1f}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom_4_z_0.9_b200_w7-renyi.png}
    \end{subfigure}
     \hspace{0.5em}
    \begin{subfigure}{0.23\textwidth}
    \caption{Binary map  $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:sim1g}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom_4_z_0.85_b200_w7-tsallis.png}
        \end{subfigure}
    \hspace{0.5em}
    \begin{subfigure}{0.23\textwidth}
    \caption{Binary~map~$\small{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
        \label{fig:sim1h}
        \includegraphics[width=\linewidth]{Figures/H_005_L5_Tsallis_adaptive_085.png}
    \end{subfigure}
   %\caption*{\footnotesize\textbf{Source:} The author (2025)}
\end{figure}
\vspace{-0.25\baselineskip} 
\begin{figure}[hbtp]
    \centering
 \caption{Results on simulated data ($L = 9$) using $S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$, and $S_{\widetilde{T}_{\lambda}}$. (a-c) $p$-value maps with a fixed $7\times7$ window.
(d) $p$-value map with adaptive windows ($W_{\min}=5$, $W_{\max}=11$) for
$\eta=3.0$. (e-h) Binary decision maps at 0.05 significance level.}
    \label{fig:sim2t}
    % Primera fila
    \begin{subfigure}{0.24\textwidth}
    \caption{$p$-value map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:sim2a}
        \includegraphics[width=\linewidth]{Figures/p-values_L9_1_8_b100_w7-AO.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:sim2b}
        \includegraphics[width=\linewidth]{Figures/p-values_L9_SIM_renyi_7x7.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
    \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:sim2c}
        \includegraphics[width=\linewidth]{Figures/p-values_L9_SIM_tsallis_7x7.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
    \caption{$p$-value map $\tiny{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
        \label{fig:sim2d}
        \includegraphics[width=\linewidth]{Figures/p-values_L9_SIM_tsallis_adaptative.png}
            \end{subfigure}
    \vspace{1em} % Espacio entre filas

    % Segunda fila
    \hspace{-0.5em}
    \begin{subfigure}{0.23\textwidth}
    \caption{Binary map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:sim2e}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom4_L9_1_8_b100_w7-AO.png}
            \end{subfigure}
    \hspace{0.5em}
    \begin{subfigure}{0.23\textwidth}
    \caption{Binary map  $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:sim2f}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom4_L9_1_8_b100_w7-09-renyi.png}
    \end{subfigure}
     \hspace{0.5em}
    \begin{subfigure}{0.23\textwidth}
    \caption{Binary map  $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:sim2g}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom4_L9_1_8_0.85_b100_w7-tsallis.png}
        \end{subfigure}
    \hspace{0.5em}
    \begin{subfigure}{0.23\textwidth}
    \caption{Binary~map~$\small{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
        \label{fig:sim2h}
        \includegraphics[width=\linewidth]{Figures/H_005_L9_tsallis_adaptative.png}
    \end{subfigure}
   %\caption*{\footnotesize\textbf{Source:} The author (2025)}
\end{figure}
\endgroup
```

\newpage

@@

### SAR Dataset and Preprocessing {#sec:SAR-data}

We evaluated the proposed test statistics using a set of five SAR images acquired from different radar missions. These images cover diverse locations: the surroundings of London (United Kingdom), the outskirts of Munich, the city of New Orleans (USA), Dublin Port (Ireland), and the coast of Jalisco (Mexico), as shown in Figures&nbsp;\ref{fig:london-sar}–\ref{fig:ja-sar}. The selected scenes include a variety of land types, such as urban areas, mountainous regions, agricultural zones, and water bodies. This diversity allows us to assess the performance of the proposed statistical tests under different texture conditions.

To facilitate interpretation, we also included the corresponding optical images (Figures&nbsp;\ref{fig:london-optical}–\ref{fig:ja-optical}) obtained from Sentinel-2 Level-2A products. These optical scenes were acquired on dates close to those of the SAR acquisitions to ensure visual consistency. The optical data were downloaded from the Copernicus Data Space platform\footnote{\url{https://dataspace.copernicus.eu/}}.

The SAR images come from TanDEM-X (X-band), UAVSAR (L-band), and Sentinel-1B (C-band). The two UAVSAR images were obtained from the Alaska Satellite Facility (ASF)\footnote{\url{https://search.asf.alaska.edu/}}, which provides airborne and satellite SAR data. The X-band images were downloaded from ESA’s Third Party Missions (ESA TPM)\footnote{\url{https://tpm-ds.eo.esa.int/collections/}}, and the C-band image from Copernicus Data Space. All images use single polarization (HH or VV), and their acquisition settings differ in resolution, scene size, and number of looks. Table&nbsp;\ref{tab:table_param} lists the main characteristics of each SAR image.

@@

SAR_data <- data.frame(
  Site = c("London",  "Munich","New Orleans", "Dublin","Coast of Jalisco" ),
  Mission = c("TanDEM-X", "UAVSAR", "UAVSAR", "TanDEM-X", "Sentinel-1B"),
  Band = c("X", "L","L", "X", "C"),
  Polarization = c("HH", "HH","HH","HH", "VV"),
  Size = c("$2000\\times2000$",  "$1024\\times1024$","$1400\\times1400$", "$1100\\times1100$","$512\\times512$"),
  L = c(1,  12, 12,  16,18),
  Resol = c("$0.99/0.99$",  "$4.9/7.2$","$4.9/7.2$", "$1.35/1.35$","$40/40$"),
  Date = c("12-11-2021", "16-04-2015","13-03-2016", "03-09-2017","29-08-2021")
)

@@

We applied the three bootstrap statistics, $S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$, and $S_{\widetilde{T}_{\lambda}}$, to the five SAR intensity images in Figures&nbsp;\ref{fig:london-sar}–\ref{fig:ja-sar}. The fixed approach uses local sliding windows of size $7\times7$. The adaptive approach (used with Tsallis) varies the window side from $5\times5$ ($N_{\min}=5$) to $11\times11$ ($N_{\max}=11$), with smoothing parameter $\eta=3$, and $B=100$ bootstrap replications per window. Each figure is organized in three rows: (i) SAR scene and a matching Sentinel-2 optical image, (ii) $p$-value maps for the three tests, and (iii) binary maps at the $5\%$ level (black = heterogeneous, white = homogeneous).

Across all scenes, the three tests detect the main textured elements (urban blocks, port facilities, roads, crop boundaries, mountains, and river banks). In the $p$-value maps, dark red ($p\!\approx\!0$) concentrates on heterogeneous areas such as urban blocks, harbor structures, and field edges, while blue ($p\!\approx\!1$) dominates open water and other uniform surfaces. Intermediate colors mark transitional textures.

The adaptive Tsallis results are visually better than the fixed windows: in homogeneous regions, the maps look smoother and more stable, while in heterogeneous regions the edges are sharper and fine details are preserved. The binary maps make this clear: the adaptive strategy keeps thin structures (roads, levees, dock lines) that are partly lost with the fixed window.

The number of looks also matters. With fewer looks (stronger speckle), the adaptive scheme improves urban and boundary detection compared with the fixed window. As $L$ increases, speckle decreases and all methods improve, yet the adaptive maps still show cleaner homogeneous areas and crisper boundaries in textured zones.

Comparing statistics, the Rényi-based $S_{\widetilde{R}_{\lambda}}$ is more sensitive to texture changes and often highlights fine-scale heterogeneity. The Tsallis-based $S_{\widetilde{T}_{\lambda}}$ is more conservative in uniform areas, reducing false alarms and producing cleaner homogeneous regions. The adaptive Tsallis variant combines these benefits with data-driven window sizes, which explains its overall better visual quality across the five scenes.

@@

```{=latex}
\begin{figure}[H]
    \centering
\caption{Detection of heterogeneous areas in a SAR image over London, UK: comparison of test statistics $\small{S_{\widetilde{H}_{\text{AO}}}}$, $\small{S_{\widetilde{R}_{\lambda}}}$, and $\small{S_{\widetilde{T}_{\lambda}}}$. Top: (a-b) input images (SAR and optical). Middle: (c-e) $p$-value maps with a fixed $7\times7$ window.
(f) $p$-value map with adaptive windows ($W_{\min}=5$, $W_{\max}=11$) for
$\eta=3.0$. Bottom:  (g-j) Binary decision maps at 0.05 significance level.}
    \label{fig:london}
    % Primera fila: 2 imágenes
    \begin{subfigure}{0.28\textwidth}
            \caption{SAR image}
        \label{fig:london-sar}
        \includegraphics[width=\linewidth]{Figures/London-sar.jpg}
    \end{subfigure}
    \hspace{2em}
    \begin{subfigure}{0.28\textwidth}
            \caption{Optical image}
        \label{fig:london-optical}
        \includegraphics[width=\linewidth]{Figures/london-optical.jpg}
    \end{subfigure}

    \vspace{1.8em}
